\section{General sets and bags}
\redno{5} What general set and bag library should R7RS-large provide?

\begin{quote}
SRFI 113 provides linear-update sets and bags, which can be mutable or
immutable according to the implementer's preference. Since the interface
of SRFI 113 depends on SRFI 114, which has been replaced by SRFI 128,
voting for SRFI 113 will cause your vote on comparator libraries to be
automatically changed to SRFI 128.

\TODO{Incorporate these post-finalization notes into the text, if not
  already done. Maybe
  comparators should come before this SRFI?}

\textbf{Post-finalization note 1}: Because SRFI 114 has been deprecated
by SRFI 128, it is recommended that implementers make use of SRFI 128
rather than SRFI 114 comparators where comparators are specified in this
SRFI. Specifically, the procedures \texttt{set}, \texttt{bag},
\texttt{set-unfold}, \texttt{bag-unfold}, \texttt{set-map},
\texttt{list-\textgreater{}set}, \texttt{list-\textgreater{}bag}, and
\texttt{alist-\textgreater{}bag}, should accept SRFI 128 rather than
SRFI 114 comparators as arguments. By the same token, the results of
\texttt{set-element-comparator} and \texttt{bag-element-comparator}, as
well as the values of \texttt{set-comparator} and
\texttt{bag-comparator}, should be SRFI 128 comparators. The sample
implementation has been updated to depend on SRFI 128 rather than SRFI
114.

\textbf{Post-finalization note 2}: The order of arguments of
\texttt{set-map} and \texttt{set-unfold} are not consistent with those
of \texttt{hash-table-map} and \texttt{hash-table-unfold} in
\href{https://srfi.schemers.org/srfi-125/}{SRFI 125}. Both SRFI 113 and
SRFI 125 are going to be part of R7RS large (also known as the
\href{http://trac.sacrideo.us/wg/wiki/RedEdition}{Red Edition}), and the
plan is to make them consistent there, following the order in SRFI 125.
Furthermore, \href{https://srfi.schemers.org/srfi-146/}{SRFI 146} (in
draft status at the time of this writing) uses the order of SRFI 125.
Since this problem was discovered after SRFI 113 was finalized, and it
wasn't an error in the SRFI, it's too late to fix it here. However, John
Cowan, the author, encourages implementers to consider adopting the
order of SRFI 125. Thanks to Marc Nieper-Wißkirchen for reporting this
mismatch, and for providing a version of SRFI 113 with the recommended
argument order in the implementation of SRFI 146.
\end{quote}


\emph{Sets} and \emph{bags} (also known as multisets) are unordered
collections that can contain any Scheme object. Sets enforce the
constraint that no two elements can be the same in the sense of the
set's associated \emph{equality predicate}; bags do not.

\setlibname{set}


Sets and bags are mutually disjoint, and disjoint from other types of
Scheme objects.

It is an error for any procedure defined in this SRFI to be invoked on
sets or bags with distinct comparators (in the sense of \texttt{eq?}).

It is an error to mutate any object while it is contained in a set or
bag.

It is an error to add an object to a set or bag which does not satisfy
the type test predicate of the comparator.

It is an error to add or remove an object for a set or a bag while
iterating over it.

The procedures of this SRFI, by default, are ``pure functional'':
they do not alter their parameters. However, this SRFI also defines
``linear-update'' procedures, all of whose names end in \texttt{!}. They
have hybrid pure-functional/side-effecting semantics: they are allowed,
but not required, to side-effect one of their parameters in order to
construct their result. An implementation may legally implement these
procedures as pure, side-effect-free functions, or it may implement them
using side effects, depending upon the details of what is the most
efficient or simple to implement in terms of the underlying
representation.

It is an error to rely upon these procedures working by side effect. For
example, this is not guaranteed to work:

\begin{verbatim}
        (let* ((set1 (set 'a 'b 'c))      ; set1 = {a,b,c}.
               (set2 (set-adjoin! set1 'd)))   ; Add d to {a,b,c}.
          set1) ; Could be either {a,b,c} or {a,b,c,d}.
\end{verbatim}

However, this is well-defined:

\begin{verbatim}
        (let ((set1 (set 'a 'b 'c)))
          (set-adjoin! set1 'd)) ; Add d to {a,b,c}.
\end{verbatim}

So clients of these procedures write in a functional style, but must
additionally be sure that, when the procedure is called, there are no
other live pointers to the potentially-modified set or bag (hence the
term ``linear update'').

There are two benefits to this convention:

\begin{itemize}
\item
  Implementations are free to provide the most efficient possible
  implementation, either functional or side-effecting.
\item
  Programmers may nonetheless continue to assume that sets are purely
  functional data structures: they may be reliably shared without
  needing to be copied, uniquified, and so forth.
\end{itemize}

In practice, these procedures are most useful for efficiently
constructing sets and bags in a side-effecting manner, in some limited
local context, before passing the set or bag outside the local
construction scope to be used in a functional manner.

Scheme provides no assistance in checking the linearity of the
potentially side-effected parameters passed to these functions ---
there's no linear type checker or run-time mechanism for detecting
violations.

Note that if an implementation uses no side effects at all, it is
allowed to return existing sets and bags rather than newly allocated
ones, even where this SRFI explicitly says otherwise.

\subsection{Comparator restrictions}

Implementations of this SRFI are allowed to place restrictions on the
comparators that the procedures accept. In particular, an implementation
may require comparators to provide a comparison procedure.
Alternatively, an implementation may require comparators to provide a
hash function, unless the equality predicate of the comparator is
\texttt{eq?}, \texttt{eqv?}, \texttt{equal?}, \texttt{string=?}, or
\texttt{string-ci=?}. Implementations must not require the provision of
both a comparison procedure and a hash function.

\subsection{Index}\label{Index}

\begin{itemize}
\item
  \protect\hyperlink{Constructors}{Constructors}: \texttt{set},
  \texttt{set-contains?}, \texttt{set-unfold}
\item
  \protect\hyperlink{Predicates}{Predicates}: \texttt{set?},
  \texttt{set-empty?}, \texttt{set-disjoint?}
\item
  \protect\hyperlink{Accessors}{Accessors}: \texttt{set-member},
  \texttt{set-element-comparator}
\item
  \protect\hyperlink{Updaters}{Updaters}: \texttt{set-adjoin},
  \texttt{set-adjoin!}, \texttt{set-replace}, \texttt{set-replace!},
  \texttt{set-delete}, \texttt{set-delete!}, \texttt{set-delete-all},
  \texttt{set-delete-all!}, \texttt{set-search!}
\item
  \protect\hyperlink{Thewholeset}{The whole set}: \texttt{set-size},
  \texttt{set-find}, \texttt{set-count}, \texttt{set-any?},
  \texttt{set-every?}
\item
  \protect\hyperlink{Mappingandfolding}{Mapping and folding}:
  \texttt{set-map}, \texttt{set-for-each}, \texttt{set-fold},
  \texttt{set-filter}, \texttt{set-filter!}, \texttt{set-remove},
  \texttt{set-remove!}, \texttt{set-partition}, \texttt{set-partition!}
\item
  \protect\hyperlink{Copyingandconversion}{Copying and conversion}:
  \texttt{set-copy}, \texttt{set-\textgreater{}list},
  \texttt{list-\textgreater{}set}, \texttt{list-\textgreater{}set!}
\item
  \protect\hyperlink{Subsets}{Subsets}: \texttt{set=?},
  \texttt{set\textless{}?}, \texttt{set\textgreater{}?},
  \texttt{set\textless{}=?}, \texttt{set\textgreater{}=?}
\item
  \protect\hyperlink{Settheoryoperations}{Set theory operations}:
  \texttt{set-union}, \texttt{set-intersection},
  \texttt{set-difference}, \texttt{set-xor}, \texttt{set-union!},
  \texttt{set-intersection!}, \texttt{set-difference!},
  \texttt{set-xor!}
\item
  \protect\hyperlink{Additionalbagprocedures}{Additional bag
  procedures}: \texttt{bag-sum}, \texttt{bag-sum!},
  \texttt{bag-product}, \texttt{bag-product!},
  \texttt{bag-element-count}, \texttt{bag-for-each-unique},
  \texttt{bag-fold-unique}, \texttt{bag-increment!},
  \texttt{bag-decrement!}, \texttt{bag-\textgreater{}set},
  \texttt{set-\textgreater{}bag}, \texttt{set-\textgreater{}bag!}
\item
  \protect\hyperlink{Comparators}{Comparators}: \texttt{set-comparator},
  \texttt{bag-comparator}
\end{itemize}

\subsection{Set procedures}

\begin{entry}{%
  \Proto{set}{ comparator element \ldots}{procedure}{set}}

  Returns a newly allocated empty set. The \emph{comparator} argument
  is a \href{http://srfi.schemers.org/srfi-114/srfi-114.html}{SRFI
    114} comparator, which is used to control and distinguish the
  elements of the set. The \emph{elements} are used to initialize the
  set.
\end{entry}

\begin{entry}{%
  \Proto{set-unfold}{ comparator stop? mapper successor
    seed}{procedure}{set}}

  Create a newly allocated set as if by \texttt{set} using
  \emph{comparator}. If the result of applying the predicate
  \emph{stop?}  to \emph{seed} is true, return the set. Otherwise,
  apply the procedure \emph{mapper} to \emph{seed}. The value that
  \emph{mapper} returns is added to the set. Then get a new seed by
  applying the procedure \emph{successor} to \emph{seed}, and repeat
  this algorithm.
\end{entry}

\subsubsection{Predicates}

\begin{entry}{%
  \Proto{set?}{ obj}{procedure}{boolean}}

  Returns \texttt{\#t} if \emph{obj} is a set, and \texttt{\#f}
  otherwise.
\end{entry}

\begin{entry}{%
  \Proto{set-contains?}{ set element}{procedure}{boolean}}

  Returns \texttt{\#t} if \emph{element} is a member of \emph{set} and
  \texttt{\#f} otherwise.
\end{entry}

\begin{entry}{%
  \Proto{set-empty?}{ set}{procedure}{boolean}}

  Returns \texttt{\#t} if \emph{set} has no elements and \texttt{\#f}
  otherwise.
\end{entry}


\begin{entry}{%
  \Proto{set-disjoint?}{ set$_1$ set$_2$}{procedure}{boolean}}

  Returns \texttt{\#t} if \emph{set$_1$} and
  \emph{set$_2$} have no elements in common and
  \texttt{\#f} otherwise.
\end{entry}

\subsubsection{Accessors}

\begin{entry}{%
  \Proto{set-member}{ set element default}{procedure}{value???}}

  Returns the element of \emph{set} that is equal, in the sense of
  \emph{set's} equality predicate, to \emph{element}. If
  \emph{element} is not a member of \emph{set}, \emph{default} is
  returned.
\end{entry}

\begin{entry}{%
  \Proto{set-element-comparator}{ set}{procedure}{comparator}}

  Returns the comparator used to compare the elements of \emph{set}.
\end{entry}

\subsubsection{Updaters}

\begin{entry}{%
  \Proto{set-adjoin}{ set element \ldots}{procedure}{set???}}

  The \texttt{set-adjoin} procedure returns a newly allocated set that
  uses the same comparator as \emph{set} and contains all the values
  of \emph{set}, and in addition each \emph{element} unless it is
  already equal (in the sense of the comparator) to one of the
  existing or newly added members. It is an error to add an element to
  \emph{set} that does not return \texttt{\#t} when passed to the type
  test procedure of the comparator.
\end{entry}

\begin{entry}{%
  \Proto{set-adjoin!}{ set element \ldots}{procedure}{set???}}

  The \texttt{set-adjoin!} procedure is the same as
  \texttt{set-adjoin}, except that it is permitted to mutate and
  return the \emph{set} argument rather than allocating a new set.
\end{entry}
\
\begin{entry}{%
  \Proto{set-replace}{ set element}{procedure}{set???}}

  The \texttt{set-replace} procedure returns a newly allocated set
  that uses the same comparator as \emph{set} and contains all the
  values of \emph{set} except as follows: If \emph{element} is equal
  (in the sense of \emph{set}'s comparator) to an existing member of
  \emph{set}, then that member is omitted and replaced by
  \emph{element}. If there is no such element in \emph{set}, then
  \emph{set} is returned unchanged.
\end{entry}

\begin{entry}{%
  \Proto{set-replace!}{ set element}{procedure}{set???}}

  The \texttt{set-replace!} procedure is the same
  as\texttt{set-replace}, except that it is permitted to mutate and
  return the \emph{set} argument rather than allocating a new set.
\end{entry}

\begin{entry}{%
  \Proto{set-delete}{ set element \ldots}{procedure}{set???}
  \Proto{set-delete!}{ set element \ldots}{procedure}{set???}
  \Proto{set-delete-all}{ set element-list}{procedure}{set???}
  \Proto{set-delete-all!}{ set element-list}{procedure}{set???}}

  The \texttt{set-delete} procedure returns a newly allocated set
  containing all the values of \emph{set} except for any that are
  equal (in the sense of \emph{set}'s comparator) to one or more of
  the \emph{elements}. Any \emph{element} that is not equal to some
  member of the set is ignored.

  The \texttt{set-delete!} procedure is the same as
  \texttt{set-delete}, except that it is permitted to mutate and
  return the \emph{set} argument rather than allocating a new set.

  The \texttt{set-delete-all} and \texttt{set-delete-all!} procedures
  are the same as \texttt{set-delete} and \texttt{set-delete!}, except
  that they accept a single argument which is a list of elements to be
  deleted.
\end{entry}

\begin{entry}{%
  \Proto{set-search!}{ set element failure success}{procedure}{set??? obj??}}

  The \emph{set} is searched for \emph{element}. If it is not found,
  then the \emph{failure} procedure is tail-called with two
  continuation arguments, \emph{insert} and \emph{ignore}, and is
  expected to tail-call one of them. If \emph{element} is found, then
  the \emph{success} procedure is tail-called with the matching
  element of \emph{set} and two continuations, \emph{update} and
  \emph{remove}, and is expected to tail-call one of them.

  The effects of the continuations are as follows (where \emph{obj} is
  any Scheme object):

  \begin{itemize}
  \item Invoking \texttt{(}\emph{insert obj}\texttt{)} causes
    \emph{element} to be inserted into \emph{set}.
  \item Invoking \texttt{(}\emph{ignore obj}\texttt{)} causes
    \emph{set} to remain unchanged.
  \item Invoking \texttt{(}\emph{update new-element obj}\texttt{)}
    causes \emph{new-element} to be inserted into \emph{set} in place
    of \emph{element}.
  \item Invoking \texttt{(}\emph{remove obj}\texttt{)} causes the
    matching element of \emph{set} to be removed from it.
  \end{itemize}

  In all cases, two values are returned: the possibly updated
  \emph{set} and \emph{obj}.
\end{entry}

\subsubsection{The whole set}

\begin{entry}{%
  \Proto{set-size}{ set}{procedure}{integer}}

  Returns the number of elements in \emph{set} as an exact integer.
\end{entry}

\begin{entry}{%
  \Proto{set-find}{ predicate set failure}{procedure}{object???}}

  Returns an arbitrarily chosen element of \emph{set} that satisfies
  \emph{predicate}, or the result of invoking \emph{failure} with no
  arguments if there is none.
\end{entry}

\begin{entry}{%
  \Proto{set-count}{ predicate set}{procedure}{integer}}

  Returns the number of elements of \emph{set} that satisfy
  \emph{predicate} as an exact integer.
\end{entry}

\begin{entry}{%
  \Proto{set-any?}{ predicate set}{procedure}{boolean???}}

  Returns \texttt{\#t} if any element of \emph{set} satisfies
  \emph{predicate}, or \texttt{\#f} otherwise. Note that this differs
  from the SRFI 1 analogue because it does not return an element of
  the set.
\end{entry}

\begin{entry}{%
  \Proto{set-every?}{ predicate set}{procedure}{boolean???}}

  Returns \texttt{\#t} if every element of \emph{set} satisfies
  \emph{predicate}, or \texttt{\#f} otherwise. Note that this differs
  from the SRFI 1 analogue because it does not return an element of
  the set.
\end{entry}

\subsubsection{Mapping and folding}

\begin{entry}{%
  \Proto{set-map}{ comparator proc set}{procedure}{boolean???}}

  Applies \emph{proc} to each element of \emph{set} in arbitrary order
  and returns a newly allocated set, created as if by \texttt{(set\
  }\emph{comparator}\texttt{)}, which contains the results of the
  applications. For example:

\begin{verbatim}
(set-map string-ci-comparator 
         symbol->string (set eq? 'foo 'bar 'baz))
             => (set string-ci-comparator "foo" "bar" "baz")
\end{verbatim}

  Note that, when \emph{proc} defines a mapping that is not 1:1, some
  of the mapped objects may be equivalent in the sense of
  \emph{comparator's} equality predicate, and in this case duplicate
  elements are omitted as in the set constructor. For example:

\begin{verbatim}
(set-map (lambda (x) (quotient x 2))
         integer-comparator
         (set integer-comparator 1 2 3 4 5))
 => (set integer-comparator 0 1 2)
\end{verbatim}

  If the elements are the same in the sense of \texttt{eqv?}, it is
  unpredictable which one will be preserved in the result.
\end{entry}

\begin{entry}{%
  \Proto{set-for-each}{ proc set}{procedure}{unspecified???}}

  Applies \emph{proc} to \emph{set} in arbitrary order, discarding the
  returned values. Returns an unspecified result.
\end{entry}

\begin{entry}{%
  \Proto{set-fold}{ proc nil set}{procedure}{object???}}

  Invokes \emph{proc} on each member of \emph{set} in arbitrary order,
  passing the result of the previous invocation as a second
  argument. For the first invocation, \emph{nil} is used as the second
  argument. Returns the result of the last invocation, or \emph{nil}
  if there was no invocation.
\end{entry}

\begin{entry}{%
  \Proto{set-filter}{ predicate set}{procedure}{set???}}

  Returns a newly allocated set with the same comparator as
  \emph{set}, containing just the elements of \emph{set} that satisfy
  \emph{predicate}.
\end{entry}

\begin{entry}{%
  \Proto{set-filter!}{ predicate set}{procedure}{set???}}

  A linear update procedure that returns a set containing just the
  elements of \emph{set} that satisfy \emph{predicate}.
\end{entry}

\begin{entry}{%
  \Proto{set-remove}{ predicate set}{procedure}{set???}}

  Returns a newly allocated set with the same comparator as
  \emph{set}, containing just the elements of \emph{set} that do not
  satisfy \emph{predicate}.
\end{entry}

\begin{entry}{%
  \Proto{set-remove!}{ predicate set}{procedure}{set???}}

  A linear update procedure that returns a set containing just the
  elements of \emph{set} that do not satisfy \emph{predicate}.
\end{entry}

\begin{entry}{%
  \Proto{set-partition}{ predicate set}{procedure}{set??? set???}}

  Returns two values: a newly allocated set with the same comparator
  as \emph{set} that contains just the elements of \emph{set} that
  satisfy \emph{predicate}, and another newly allocated set, also with
  the same comparator, that contains just the elements of \emph{set}
  that do not satisfy \emph{predicate}.
\end{entry}

\begin{entry}{%
  \Proto{set-partition!}{ predicate set}{procedure}{set??? set???}}

  A linear update procedure that returns two sets containing the
  elements of \emph{set} that do and do not, respectively, not satisfy
  \emph{predicate}.
\end{entry}

\subsubsection{Copying and conversion}

\begin{entry}{%
  \Proto{set-copy}{ set}{procedure}{set???}}

  Returns a newly allocated set containing the elements of \emph{set},
  and using the same comparator.
\end{entry}

\begin{entry}{
  \Proto{set->list}{ set}{procedure}{list???}}

  Returns a newly allocated list containing the members of \emph{set}
  in unspecified order.
\end{entry}

\begin{entry}{%
  \Proto{list->set}{ comparator list}{procedure}{set???}}

  Returns a newly allocated set, created as if by \texttt{set} using
  \emph{comparator}, that contains the elements of
  \emph{list}. Duplicate elements (in the sense of the equality
  predicate) are omitted.
\end{entry}

\begin{entry}{%
  \texttt{(list->set!}{ set list}{procedure}{set???}}

  Returns a set that contains the elements of both \emph{set} and
  \emph{list}. Duplicate elements (in the sense of the equality
  predicate) are omitted.
\end{entry}

\subsubsection{Subsets}

Note: The following three predicates do not obey the trichotomy law and
therefore do not constitute a total order on sets.

\TODO{Explain how elements are compared. Via their comparators?}

\begin{entry}{%
  \Proto{set=?}{ set$_1$ set$_2$ \ldots}{procedure}{boolean???}}

  Returns \texttt{\#t} if each \emph{set} contains the same elements.
\end{entry}

\begin{entry}{%
  \Proto{set<?}\emph{set$_1$ set$_2$ \ldots}{procedure}{boolean???}}

Returns \texttt{\#t} if each \emph{set} other than the last is a
proper subset of the following \emph{set}, and \texttt{\#f} otherwise.
\end{entry}

\begin{entry}{%
  \Proto{set>?}{set$_1$ set$_2$ \ldots}{procedure}{boolean???}}

  Returns \texttt{\#t} if each \emph{set} other than the last is a
  proper superset of the following \emph{set}, and \texttt{\#f}
  otherwise.
\end{entry}

\begin{entry}{%
  \Proto{set<=?}{ set$_1$ set$_2$ \ldots}{procedure}{boolean???}}

  Returns \texttt{\#t} if each \emph{set} other than the last is a
  subset of the following \emph{set}, and \texttt{\#f} otherwise.
\end{entry}

\begin{entry}{%
  \Proto{set>=?}{ set$_1$ set$_2$ \ldots}{procedure}{boolean???}}

  Returns \texttt{\#t} if each \emph{set} other than the last is a
  superset of the following \emph{set}, and \texttt{\#f} otherwise.
\end{entry}

\subsubsection{Set theory operations}

\begin{entry}{%
  \Proto{set-union}{ set$_1$ set$_2$ \ldots}{procedure}{set???}
  \Proto{set-intersection}{ set$_1$ set$_2$ \ldots}{procedure}{set???}
  \Proto{set-difference}{ set$_1$ set$_2$ \ldots}{procedure}{set???}
  \Proto{set-xor}{ set$_1$ set$_2$}{procedure}{set???}}

  Return a newly allocated set that is the union, intersection,
  asymmetric difference, or symmetric difference of the
  \emph{sets}. Asymmetric difference is extended to more than two sets
  by taking the difference between the first set and the union of the
  others. Symmetric difference is not extended beyond two
  sets. Elements in the result set are drawn from the first set in
  which they appear.
\end{entry}

\begin{entry}{%
  \Proto{set-union!}{ set$_1$ set$_2$ \ldots}{procedure}{set???}
  \Proto{set-intersection!}{ set$_1$ set$_2$ \ldots}{procedure}{set???}
  \Proto{set-difference!}{ set$_1$ set$_2$ \ldots}{procedure}{set???}
  \Proto{set-xor! }{ set$_1$ set$_2$}{procedure}{set???}}

  Linear update procedures returning a set that is the union,
  intersection, asymmetric difference, or symmetric difference of the
  \emph{sets}. Asymmetric difference is extended to more than two sets
  by taking the difference between the first set and the union of the
  others.  Symmetric difference is not extended beyond two
  sets. Elements in the result set are drawn from the first set in
  which they appear.
\end{entry}

\subsection{Bag procedures}\label{Bagprocedures}

Bags are like sets, but can contain the same object more than once.
However, if two elements that are the same in the sense of the equality
predicate, but not in the sense of \texttt{eqv?}, are both included, it
is not guaranteed that they will remain distinct when retrieved from the
bag. It is an error for a single procedure to be invoked on bags with
different comparators.

\TODO{This must be rewritten in standard Scheme entry format.}

The procedures for creating and manipulating bags are the same as those
for sets, except that \texttt{set} is replaced by \texttt{bag} in their
names, and that adjoining an element to a bag is effective even if the
bag already contains the element. If two elements in a bag are the same
in the sense of the bag's comparator, the implementation may in fact
store just one of them.

The \texttt{bag-union}, \texttt{bag-intersection},
\texttt{bag-difference}, and \texttt{bag-xor} procedures (and their
linear update analogues) behave as follows when both bags contain
elements that are equal in the sense of the bags' comparator:

\begin{itemize}
\item
  For \texttt{bag-union}, the number of equal elements in the result is
  the largest number of equal elements in any of the original bags.
\item
  For \texttt{bag-intersection}, the number of equal elements in the
  result is the smallest number of equal elements in any of the original
  bags.
\item
  For \texttt{bag-difference}, the number of equal elements in the
  result is the number of equal elements in the first bag, minus the
  number of elements in the other bags (but not less than zero).
\item
  For \texttt{bag-xor}, the number of equal elements in the result is
  the absolute value of the difference between the number of equal
  elements in the first and second bags.
\end{itemize}

\subsubsection{Additional bag procedures}

\begin{entry}{%
  \Proto{bag-sum}{ set$_1$ set$_2$ \ldots}{procedure}{bag???}
  \Proto{bag-sum!}{ bag$_1$ bag$_2$ \ldots}{procedure}{bag???}}

  \TODO{are those really \texttt{set} arguments in \texttt{bag-sum}?}

  The \texttt{bag-sum} procedure returns a newly allocated bag
  containing all the unique elements in all the \emph{bags}, such that
  the count of each unique element in the result is equal to the sum
  of the counts of that element in the arguments. It differs from
  \texttt{bag-union} by treating identical elements as potentially
  distinct rather than attempting to match them up.

  The \texttt{bag-sum!} procedure is equivalent except that it is
  linear-update.
\end{entry}

\begin{entry}{%
  \Proto{bag-product}{ n bag}{procedure}{bag???}
  \Proto{bag-product!}{ n bag}{procedure}{bag???}}

  The \texttt{bag-product} procedure returns a newly allocated bag
  containing all the unique elements in \emph{bag}, where the count of
  each unique element in the bag is equal to the count of that element
  in \texttt{bag} multiplied by \emph{n}.

  The \texttt{bag-product!} procedure is equivalent except that it is
  linear-update.
\end{entry}

\begin{entry}{%
  \Proto{bag-unique-size}{ bag}{procedure}{integer???}}

  Returns the number of unique elements of \emph{bag} \TODO{as an
    exact integer???}.
\end{entry}

\begin{entry}{%
  \Proto{bag-element-count}{ bag element}{procedure}{integer???}}

  Returns an exact integer representing the number of times that
  \emph{element} appears in \emph{bag}.
\end{entry}

\begin{entry}{%
  \Proto{bag-for-each-unique}{ proc bag}{procedure}{unspecified???}}

  Applies \emph{proc} to each unique element of \emph{bag} in
  arbitrary order, passing the element and the number of times it
  occurs in \emph{bag}, and discarding the returned values. Returns an
  unspecified result.
\end{entry}

\begin{entry}{%
  \Proto{bag-fold-unique}{ proc nil bag}{procedure}{value???}}

  Invokes \emph{proc} on each unique element of \emph{bag} in
  arbitrary order, passing the number of occurrences as a second
  argument and the result of the previous invocation as a third
  argument. For the first invocation, \emph{nil} is used as the third
  argument. Returns the result of the last invocation.
\end{entry}

\begin{entry}{%
  \Proto{bag-increment!}{ bag element count}{procedure}{bag???}
  \Proto{bag-decrement!}{ bag element count}{procedure}{bag???}}

  Linear update procedures that return a bag with the same elements as
  \emph{bag}, but with the element count of \emph{element} in
  \emph{bag} increased or decreased by the exact integer \emph{count}
  (but not less than zero).
\end{entry}

\begin{entry}{%
  \Proto{bag->set}{ bag}{procedure}{set???}
  \Proto{set->bag}{ set}{procedure}{bag???}
  \Proto{set->bag!}{ bag set}{procedure}{bag???}}

  The \texttt{bag-\textgreater{}set} procedure returns a newly
  allocated set containing the unique elements (in the sense of the
  equality predicate) of \emph{bag}. The
  \texttt{set-\textgreater{}bag} procedure returns a newly allocated
  bag containing the elements of \emph{set}. The
  \texttt{set-\textgreater{}bag!} procedure returns a bag containing
  the elements of both \emph{bag} and \emph{set}. In all cases, the
  comparator of the result is the same as the comparator of the
  argument or arguments.
\end{entry}

\begin{entry}{%
  \Proto{bag->alist}{ bag}{procedure}{alist???}
  \Proto{alist->bag}{ comparator alist}{procedure}{bag???}}

  The \texttt{bag-\textgreater{}alist} procedure returns a newly
  allocated alist whose keys are the unique elements of \emph{bag} and
  whose values are the number of occurrences of each element. The
  \texttt{alist-\textgreater{}bag} returning a newly allocated bag
  based on \emph{comparator}, where the keys of \emph{alist} specify
  the elements and the corresponding values of \emph{alist} specify
  how many times they occur.
\end{entry}

\subsection{Comparators}

\TODO{This section must be expanded.}

The following comparators are used to compare sets or bags, and allow
sets of sets, bags of sets, etc.

\texttt{set-comparator}

\texttt{bag-comparator}

Note that these comparators do not provide comparison procedures, as
there is no ordering between sets or bags. It is an error to compare
sets or bags with different element comparators.


Editor:
\href{mailto:srfi-editors\%20at\%20srfi\%20dot\%20schemers\%20dot\%20org}{Mike
Sperber}
