\section{{vector library Rationale}}

R5RS provides very few list-processing
procedures, for which reason \protect\hyperlink{SRFI-1}{SRFI 1} exists.
However, R5RS provides even fewer vector
operations --- while it provides mapping, appending, et cetera
operations for lists, it specifies only nine vector manipulation
operations:

\begin{itemize}
\tightlist
\item
  \texttt{vector?}
\item
  \texttt{make-vector}
\item
  \texttt{vector}
\item
  \texttt{vector-length}
\item
  \texttt{vector-ref}
\item
  \texttt{vector-set!}
\item
  \texttt{vector-\textgreater{}list}
\item
  \texttt{list-\textgreater{}vector}
\item
  \texttt{vector-fill!}
\end{itemize}

\protect\hyperlink{R7RS-small}{R7RS-small} added support for
\emph{start} and \emph{end} arguments to
\texttt{vector-\textgreater{}list}, \texttt{list-\textgreater{}vector}
and \texttt{vector-fill!}. It also provided seven additional vector
procedures, bringing vectors and lists to approximate parity:

\begin{itemize}
\tightlist
\item
  \texttt{vector-copy}
\item
  \texttt{vector-copy!}
\item
  \texttt{vector-append}
\item
  \texttt{vector-map}
\item
  \texttt{vector-for-each}
\item
  \texttt{vector-\textgreater{}string}
\item
  \texttt{string-\textgreater{}vector}
\end{itemize}

\protect\hyperlink{SRFI-43}{SRFI 43} standardized more vector
procedures, all of which are included in this SRFI. Unfortunately,
R7RS-small and SRFI 43 placed irreconcileable requirements on the
procedures invoked by \texttt{vector-map} and \texttt{vector-for-each}.
This SRFI resolves that issue by changing these SRFI 43 procedures as
well as \texttt{vector-map!}, \texttt{vector-fold},
\texttt{vector-fold-right}, and \texttt{vector-count} to leave out the
index argument that is passed under SRFI 43's definition.

In addition, the version of \texttt{vector-copy} in this SRFI does not
require support for a fill argument, which makes it equivalent to the
R7RS-small definition.

This SRFI also provides the following new procedures (some from Python,
some from other sources):

\begin{itemize}
\tightlist
\item
  \texttt{vector-unfold!}
\item
  \texttt{vector-unfold-right!}
\item
  \texttt{vector-append-subvectors}
\item
  \texttt{vector-cumulate}
\item
  \texttt{vector-partition}
\end{itemize}

It should be noted that no vector sorting procedures are provided by
this SRFI, because there already are several SRFIs for that purpose.

